Stateless Model Checking with Data-Race Preemption Points

### 2015-11-20 pldi-16/evaluation.tex
put subfigure labels on figure 2
clean up todos
garth's final comments pt2, all except the tough ones

We chose P2s and Pintoses for our test suite because of the relative ease of generating hundreds of unique state spaces,
varied in size and correctness, and with a diverse set of bug types\footnote{
Many of the codebases exhibited {\em deterministic} bugs (i.e., encountered on the first interleaving tested).
We fixed these by hand, before running tests, to ensure that every bug in our study required meaningful work by the MC.}.
% TODO: Is this ok? Too arrogant?
We believe that merely finding a small handful of new real-world bugs is distressingly anecdotal,
and that our test suite's size allows for a more statistically significant comparison among MC and data-race testing strategies.
%compared to anecdotally showing a small handful of new bugs that could be found in real-world programs.

@@

%\quicksand's times by 10 to convert from wall-clock time to CPU-hours (even though it sometimes falls short of 100\% parallelism).
Figure~\ref{fig:dowefindbugsfaster} plots a cumulative distribution of bugs found before a given elapsed CPU-time. % during its associated test.
We tested several configurations of \quicksand~to compare against SSS-MC from multiple angles,
indicated by each of the ``QS'' lines,
which we explain presently.

@@

@@

%Figure~\ref{fig:bargraph}
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
%In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
%even compared to the other test programs
In \mxtest, in which we stop trusting the lock implementation's correctness,
the control experiment found dramatically fewer bugs (just 1)\footnote{
	The one bug SSS-MC found was in a fully-assembler lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient to find the bug.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it often suffices to assume correctly-implemented locks,
we consider this strong evidence that low-level synchronization code must be verified with data-race PPs.

@@

Among those 69 true data-races, %which initially fit the malloc-recycle pattern,
none of them exposed a new bug when used as a PP.
This suggests that for other data-race tools,
suppressing malloc-recycle candidates may be a productive heuristic,
even if unsound without Iterative Deepening.
However, \quicksand~was able to correctly identify the 69 violations of that heuristic (among 30 distinct tests),
and fall back to classifying them with DPOR.
%which we consider a much stronger verification.

{\bf Full verification.}
For 153 of our 629 tests, \quicksand~was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}}.
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of \quicksand's total verifications before a given elapsed CPU-time.
Among these, 36 tests contained no data-race candidates whatsoever,
so the same verification could be provided by SSS-MC.
We plot these for comparison,
according to the time taken by that test's SSS-MC control experiment.
When a test can be totally verified without any data-race PPs,
\quicksand~may be slower than SSS-MC due to redundant work (\sect{\ref{sec:future}}).
Of course, using data-race PPs increases our capacity for total verification by 4.25x.

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{nondet-drs-1-v2.pdf}
	\caption{Some data-race candidates may not be identified during a single program execution.
		Using nondeterministic data races as PPs,
		\quicksand~found 189\% as many data-race bugs compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}
Figure~\ref{fig:dr-falsenegs} compares the types of data-race candidates necessary to expose each data-race bug in our test suite.
The first series represents the bugs found using PPs from single-pass data-race candidates,
% TODO: re read racefuzzer/portend paper to confirm this is true.
i.e., the state-of-the-art approach used by \cite{racefuzzer,portend}.
The second series shows all data-race bugs \quicksand~found,
which includes the former type as well as new bugs involving nondeterministic data-races.
\quicksand~found 68 data-race bugs in total, only 36 of which could be found with single-pass data-race candidates alone.
%In total, we found 189\% as many bugs by using nondeterministic data-race candidates as PPs.

@@

Note that we are not comparing how much testing time is required before identifying the data-race candidates involved in each bug.
Single-pass data races can all be found after a single program execution,
while \quicksand~may potentially take up to all 10 CPU-hours before identifying a nondeterministic data race.
However, prior work data-race tools \cite{tsan}, being not integrated with a MC,
are not intended to discover new candidates under subsequent runs.
Running a single-pass data-race tool repeatedly for 10 CPU-hours could potentially uncover some nondeterministic candidates,
but stress testing's comparative problem with achieving reliable coverage is already well-understood
\cite{chess-icb,gambit}.
Likewise, replay-based tools \cite{portend} are dependent upon the data-race detector to provide an execution trace leading to each candidate.
This experiment suggests that
such tools could benefit from a similar feedback loop as Iterative Deepening.
%i.e., discovering transitively-reachable data races while testing initial ones.
%although that would still not simultaneously be able to provide total verifications.
% TODO: read gambit paper

>>>>

We chose P2s and Pintoses for our test suite because of the relative ease of generating hundreds of unique state spaces,
varied in size and correctness, and with a diverse set of bug types\footnote{
Many of the codebases exhibited {\em deterministic} bugs (i.e., encountered on the first interleaving tested).
We fixed these by hand, before running tests, to ensure that every bug in our study required meaningful work by the MC.}.
% TODO: Is this ok? Too arrogant?
We believe that merely finding a small handful of new real-world bugs is largely anecdotal,
and that our test suite's size allows for a more statistically significant comparison among MC and data-race testing strategies.
%compared to anecdotally showing a small handful of new bugs that could be found in real-world programs.

@@

%\quicksand's times by 10 to convert from wall-clock time to CPU-hours (even though it sometimes falls short of 100\% parallelism).
Figure~\ref{fig:dowefindbugsfaster} plots a cumulative distribution of bugs found before a given elapsed CPU consumption. % during its associated test.
We tested several configurations of \quicksand~to compare against SSS-MC from multiple angles,
indicated by each of the ``QS'' lines,
which we explain presently.

@@

% TODO CAMREADY: ICB

% TODO CAMREADY: Compare QS-ETAs to QS-Random to evaluate "smaller is better" claim.

@@

%Figure~\ref{fig:bargraph}
The left half of Table~\ref{tab:drbugs}
breaks down the number and types of bugs found by each test program.
% The astute observer will note that the sum of all types of failures observed is more than the total number of data-race bugs...
%TODO CAMREADY: Run a mutex expt where "all atomic instrs" are PPs. See how many bugs are missed anyway.
% Wwe re-ran the \mxtest control experiment with \landslide~hard-coded to preempt on any atomic instruction
% (as well as on the mutex API boundaries).
% Still, this smarter configuration for SSS-MC found only 99999999999 bugs of \quicksand's 13.
%In \mxtest~in particular, the control experiment found dramatically fewer bugs (just 1),
%even compared to the other test programs
In \mxtest, in which we do not trust the lock implementation's correctness,
the control experiment found dramatically fewer bugs (just 1)\footnote{
	The one bug SSS-MC found was in a fully-assembler lock implementation. {\tt yield()}'s return value clobbered a value stored in {\tt \%eax}, which could lead to a failure after two repeated contentions. Preempting only on {\tt yield()} (in the contention loop) was sufficient to find the bug.}.
%Intuitively, this is due to our control experiment being able to preempt only on the boundaries of the API which
%Though for many applications of MC, assuming a correct lock implementation is sufficient,
Though it often suffices to assume correctly-implemented locks,
we consider this strong evidence that new low-level synchronization code must be verified with data-race PPs.

@@

Among those 69 true data-races, %which initially fit the malloc-recycle pattern,
none exposed a new bug when used as a PP.
This suggests that for other data-race tools,
suppressing malloc-recycle candidates may be a productive heuristic,
even if unsound without Iterative Deepening.
However, \quicksand~was able to correctly identify the 69 violations of that heuristic (among 30 distinct tests),
and fall back to classifying them with DPOR.
%which we consider a much stronger verification.

{\bf Full verification.}
For 153 of our 629 tests, \quicksand~was able to provide the total verification guarantee described in \sect{\ref{sec:totalverif}}.
In Figure~\ref{fig:totalverif} we plot the cumulative distribution of \quicksand's total verifications before a given elapsed CPU consumption.
Among these, 36 tests contained no data-race candidates whatsoever,
so the same verification could be provided by SSS-MC.
We plot these for comparison,
according to the time taken by that test's SSS-MC control experiment.
When a test can be totally verified without any data-race PPs,
\quicksand~may be slower than SSS-MC due to redundant work (\sect{\ref{sec:future}}).
Of course, using data-race PPs increases our capacity for total verification by 4.25x.

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{nondet-drs-1-v2.pdf}
	\caption{Some data-race candidates may not be identified during a single program execution.
		Using nondeterministic data races as PPs,
		\quicksand~found 189\% as many data-race bugs compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}
Figure~\ref{fig:dr-falsenegs} compares the types of data-race candidates necessary to expose each data-race bug in our test suite.
The first series represents the bugs found using PPs from single-pass data-race candidates,
% not entirely true, as portend could be given data-race traces from an MC,
%but they don't do it in their paper, so i feel comfortable making this claim
i.e., the state-of-the-art approach used by \cite{racefuzzer,portend}.
The second series shows all data-race bugs \quicksand~found,
which includes the former type as well as new bugs involving nondeterministic data-races.
\quicksand~found 68 data-race bugs in total, only 36 of which could be found with single-pass data-race candidates alone.
%In total, we found 189\% as many bugs by using nondeterministic data-race candidates as PPs.

@@

Note that we are not comparing how much testing time is required before identifying the data-race candidates involved in each bug.
Single-pass data races can all be found after a single program execution,
while \quicksand~may potentially take up to all 10 CPU-hours before identifying a nondeterministic data race.
However, prior work data-race tools \cite{tsan}, being not integrated with a MC,
are not intended to discover new candidates under subsequent runs.
Running a single-pass data-race tool repeatedly for 10 CPU-hours could potentially uncover some nondeterministic candidates,
but stress testing's comparative problem with achieving reliable coverage is already well-understood
\cite{chess-icb,gambit}.
Likewise, replay-based tools \cite{portend} are dependent upon the data-race detector to provide an execution trace leading to each candidate.
This experiment suggests that
such tools could benefit from a similar feedback loop as used in Iterative Deepening.
%i.e., discovering transitively-reachable data races while testing initial ones.
%although that would still not simultaneously be able to provide total verifications.
% TODO: read gambit paper

### 2015-11-20 pldi-16/intro.tex
garth's final comments pt2, all except the tough ones

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate operating systems classes at Carnegie Mellon, University of California at Berkeley, and University of Chicago.
%We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
We find many of the same bugs found by the conventional approach faster using subsets of preemption points,
and that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all.

>>>>

We evaluate \quicksand~by testing \numstudence~student thread libraries and kernels from the undergraduate operating systems classes at Carnegie Mellon University, University of California at Berkeley, and University of Chicago.
%We show many advantages over both conventional stateless model checking and single-pass data-race analysis.
We find many of the same bugs found by the conventional approach faster using subsets of preemption points,
and that data-race preemption points quickly expose many new bugs that prior model checkers could not find at all.

### 2015-11-20 pldi-16/overview.tex
garth's final comments pt2, all except the tough ones

Prior work \cite{hybriddatarace} distinguishes between {\em pure happens-before} \cite{lamport-clocks},
in which accesses with a lock release/acquire in between are not considered concurrent,
and {\em limited happens-before},
in which only blocking operations like {\tt cond\_wait} or {\tt join} enforce the order.
Compared to the former, the latter may output false positives, but avoids false negatives \cite{tsan},
in which a potential racing access pair is not reported, even though it could lead to a bug.
We use the limited happens-before analysis, accepting some false positive overhead for the sake of total verification (\sect{\ref{sec:totalverif}}).

>>>>

Prior work \cite{hybriddatarace} distinguishes between {\em pure happens-before} \cite{lamport-clocks},
in which accesses with a lock release/acquire in between are not considered concurrent,
and {\em limited happens-before},
in which only blocking operations like {\tt cond\_wait} or {\tt join} enforce the order.
Compared to the former, the latter may output false positives, but avoids false negatives \cite{tsan},
in which a potential racing access pair is not reported, even though it could lead to a bug.
We use the limited happens-before analysis, accepting some overhead from false positives for the sake of total verification (\sect{\ref{sec:totalverif}}).

### 2015-11-20 pldi-16/paper.tex
garth's final comments pt2, all except the tough ones

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

Our evaluation shows that \quicksand~achieves better bug-finding results than either single-pass data-race detection or single-state-space model checking alone.
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
We find new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.

%Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
\quicksand~is open-source and its interface can be adapted to fit any tool similar to \landslide.
We have also posted our evaluation's full data set. %and logs.
They are available at:

>>>>

We have presented Iterative Deepening and \quicksand, a new technique and tool for automating the choice of preemption points (PPs) during stateless model checking.
% , and \quicksand, a tool which implements this technique tailored to the \landslide~model checker.
\quicksand~incorporates data-race analysis to create new PPs tailored specifically to the program under test,
and automatically finds state spaces that are appropriately sized to complete in a given CPU budget.
%and manages multiple model checker instances to test many different PP subsets in a given CPU budget, even when the full state space of all PPs would be computationally intractable.

We achieve better bug-finding results than either single-pass data-race detection or single-state-space model checking alone,
%We dramatically reduce false-positive data race reports,
%%by verifying the absence of bugs in their associated state spaces
%and find bugs caused by nondeterministic data-races in alternate thread interleavings missed entirely by single-pass analysis.
%Compared to prior work,
%, we find bugs faster in smaller subset state spaces, %% not true, or at least, not by very much
%when the maximal state space is too large to explore
finding new bugs with data-race PPs that could not be exposed by preempting only on synchronization APIs.
By using \numstudence~student projects as our test suite, we also show the potential benefit as a debugging platform in educational settings.
Moreover, when all data-race PPs can be fully tested within the CPU budget, we provide a verification as strong as if every single instruction had been used as a PP.

%Although \landslide~itself is not publicly available due to its dependency on the non-free simulator \simics, \quicksand~is open-source and its interface can be adapted to fit any similar tool.
\quicksand~is open-source and its interface can be adapted to fit any tool similar to \landslide.
We have also posted our evaluation's data set. %and logs.
They are available at:

### 2015-11-20 pldi-16/related.tex
clean up todos
garth's final comments pt2, all except the tough ones

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs.
However, the resulting state space explosion demands that any such tool either
choose a small subset of instructions to consider as PPs
or be limited to very small test programs.
%However, every such prior tool we know of has serious drawbacks.
SKI \cite{ski} approaches kernel code by statically choosing a random set of instructions in advance, %offsets from the start of the test,
which is perhaps more similar to
%stress testing or
schedule fuzzing \cite{randomized-scheduler} than to exhaustive state space exploration.
%
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU \cite{rcu}, which is similar to our \mxtest~experiment,
although it requires code to be written in the PROMELA language.
%However, SPIN is stateful rather than stateless, and explicitly storing visited program states rather than using DPOR limits the size of programs that can be practically tested.
%
Inspect \cite{inspect} instruments source code by inserting wrapper calls around all accesses to potentially-shared data.
It identifies such instructions in advance with an over-approximating alias analysis,
while \landslide~\cite{landslide} traces the memory locations of accesses at runtime.
Both SPIN and Inspect fix their set of PPs in advance, so could be
%combined with \quicksand~
extended with Iterative Deepening in future work.
%so cannot check implementations directly.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan} combine the lockset and happens-before analyses into a hybrid technique, which we employ.
DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%event-driven
Android applications, using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} offers data-race techniques for kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which would allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

@@

Portend \cite{portend} is the most closely related work we have found.
Based on single-pass data race reports, it tests alternate program executions to classify candidates in a taxonomy of likely severity.
It uses symbolic execution to test input nondeterminism as well as schedule nondeterminism,
%while we explore the latter only.
and additionally reports non-failing races which nevertheless cause
%suspiciously
different program output. %, which is orthogonal to our technique.
However, Portend does not test alternate interleavings {\em in advance} of knowing any specific data races,
which is necessary to expose certain bugs (\sect{\ref{sec:eval-falseneg}}) or to provide full verification (\sect{\ref{sec:totalverif}}).
% Not really true.
%It also assumes the POSIX synchronization API, so cannot verify arbitrary synchronization algorithms such as we do with \mxtest.
Future work could combine the two approaches, using MC to produce new data-race traces for Portend to classify, or using Portend's analysis to inform \quicksand's heuristic priorities.

@@

% TODO: Read OOPSLA 2015 "SATcheck, sat-directed SMC for SC/TSO"

>>>>

{\bf Preempting at instruction granularity} is a prerequisite for using data-race PPs.
However, the resulting state space explosion demands that any such tool either
choose a small subset of instructions to consider as PPs
or be limited to very small test programs.
%However, every such prior tool we know of has serious drawbacks.
SKI \cite{ski} approaches kernel code by statically choosing a random set of instructions in advance, %offsets from the start of the test,
which is perhaps more similar to
%stress testing or
schedule fuzzing \cite{randomized-scheduler} than to exhaustive state space exploration.
%
SPIN \cite{spin} specializes in verifying synchronization primitive implementations such as RCU \cite{rcu}, which is similar to our \mxtest~experiment,
although it requires code to be written in the PROMELA language.
%However, SPIN is stateful rather than stateless, and explicitly storing visited program states rather than using DPOR limits the size of programs that can be practically tested.
%
Inspect \cite{inspect} instruments source code by inserting wrapper calls around all accesses to potentially-shared data.
It identifies such instructions in advance with an over-approximating static alias analysis,
while \landslide~\cite{landslide} traces the memory locations of accesses at runtime.
Both SPIN and Inspect fix their set of PPs in advance, so could be
%combined with \quicksand~
extended with Iterative Deepening in future work.
%so cannot check implementations directly.

@@

%Too many related projects to list have made contributions to the
Many advances have been made on the false-positive data race problem since it was first introduced in \cite{eraser}.
\cite{hybriddatarace} and \cite{tsan} combine the lockset and happens-before analyses into a hybrid technique, which we employ.
DroidRacer \cite{droidracer} and CAFA \cite{cafa} target
%event-driven
Android applications, using domain-specific heuristics (orthogonal to our method) to reduce false positives. % cut for space?
Like \landslide, DataCollider \cite{datacollider} offers data-race techniques for kernel code.
IFRit \cite{ifrit}
improves performance using an interference analysis,
which could allow future work to avoid tracing every memory access.
%although in our context, we cannot admit false negatives (\sect{\ref{sec:totalverif}}).
% No, IDGAF about pure happens before.
%FastTrack \cite{fasttrack} optimizes the performance of pure happens-

@@

Portend \cite{portend} is the most closely related work we have found.
Based on reports from single-pass data-race analysis, it tests alternate executions to classify candidates in a taxonomy of likely severity.
It uses symbolic execution to test input nondeterminism as well as schedule nondeterminism,
%while we explore the latter only.
and additionally reports non-failing races which nevertheless cause
%suspiciously
different program output. %, which is orthogonal to our technique.
However, Portend does not test alternate interleavings {\em in advance} of knowing any specific data races,
which is necessary to find certain bugs (\sect{\ref{sec:eval-falseneg}}) or to provide full verification (\sect{\ref{sec:totalverif}}).
% Not really true.
%It also assumes the POSIX synchronization API, so cannot verify arbitrary synchronization algorithms such as we do with \mxtest.
Future work could combine the two approaches, using MC to produce new data-race traces for Portend to classify, or using Portend's analysis to inform \quicksand's heuristic priorities.

@@

### 2015-11-20 pldi-16/soundness.tex
garth's final comments pt2, all except the tough ones

\newcommand\ppnext[1]{\ensuremath{\mathsf{next}(#1)}}
\newcommand\ppinstr[1]{\ensuremath{\mathsf{instr}(#1)}}
\newcommand\ppothers[1]{\ensuremath{\mathsf{others}(#1)}}
\begin{theorem}
If a bug can be exposed by any thread interleaving possible by preempting on all instructions,
Iterative Deepening will eventually test an equivalent interleaving which exposes the same bug.
\end{theorem}
\begin{proof}
The proof has two parts:
first, we show that preempting on data-racing instructions and synchronization API boundaries suffices to test all possible program behavior;
second, we show that Iterative Deepening will eventually detect all such data races.
Given a PP $p$, let $\ppnext{p}$ denote the next transition after $p$ executed by the thread which ran immediately before $p$,
let $\ppinstr{p}$ denote the first instruction of $\ppnext{p}$,
and let $\ppothers{p}$ denote the transitions by other threads between $p$ and $\ppnext{p}$.

@@

We now show by contradiction there cannot be {\em multiple} data-race PPs which must both be enabled before either data race can be identified.
Assume there does not exist a single transition $t_1 \in \mathcal{S}$ which alone can be split into $\{t_1',t_1''\}$ by a PP $q$,
such that another thread's concurrent transition $t_2$ conflicts with $t_1''$.
By the soundness of DPOR, because all $t_2$s are independent with $t_1''$, $\mathcal{S} \equiv \mathcal{S} \cup q$.
Replacing $\mathcal{S}$ with $\mathcal{S} \cup q$ in the above assumption shows that no {\em pair} of new $q$s would expose new program behaviour, and inductively, no set of $q$s of any size, which contradicts the previous paragraph.

@@

We identify a particular class of false positive data-race candidates in which the memory was recycled by {\tt malloc} between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behavior.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 7 will be flagged as a data race.
We call this a {\em malloc-recycle data race}.
To the human eye, this is obviously a false positive: reordering lines 4-7 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to an initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 6.

@@

Fortunately, when data-race detection is combined with DPOR and Iterative Deepening, pruning these false positives is sound even when such adversarial programs are considered.
This makes it unnecessary to verify such data races by actively adding more preemptions,
achieving a potentially combinatorial reduction in how many state spaces we generate.
%Intuitively, we need not worry about cases such as Figure~\ref{fig:recycle-bug} because,
%should they be true races,
%DPOR will reorder threads sufficiently for the malloc-recycle pattern to disappear.
We provide a proof sketch below.

>>>>

\newcommand\ppnext[1]{\ensuremath{\mathsf{next}(#1)}}
\newcommand\ppinstr[1]{\ensuremath{\mathsf{instr}(#1)}}
\newcommand\ppothers[1]{\ensuremath{\mathsf{others}(#1)}}
\begin{theorem}
If a bug can be exposed by any thread interleaving possible by preempting on all instructions during a specific test,
Iterative Deepening will eventually test an equivalent interleaving which exposes the same bug.
\end{theorem}
\begin{proof}
The proof has two parts:
first, we show that preempting on data-racing instructions and synchronization API boundaries suffices to test all possible program behavior;
second, we show that Iterative Deepening will eventually detect all such data races.
Given a PP $p$, let $\ppnext{p}$ denote the next transition after $p$ executed by the thread which ran immediately before $p$,
let $\ppinstr{p}$ denote the first instruction of $\ppnext{p}$,
and let $\ppothers{p}$ denote the transitions by other threads between $p$ and $\ppnext{p}$.

@@

We now show by contradiction there cannot be {\em multiple} data-race PPs which must all be enabled before either data race can be identified.
Assume there does not exist a single transition $t_1 \in \mathcal{S}$ which alone can be split into $\{t_1',t_1''\}$ by a PP $q$,
such that another thread's concurrent transition $t_2$ conflicts with $t_1''$.
By the soundness of DPOR, because all $t_2$s are independent with $t_1''$, $\mathcal{S} \equiv \mathcal{S} \cup q$.
Replacing $\mathcal{S}$ with $\mathcal{S} \cup q$ in the above assumption shows that no {\em pair} of new $q$s would expose new program behaviour, and inductively, no set of $q$s of any size, which contradicts the previous paragraph.

@@

We identify a particular class of false positive data-race candidates in which the associated memory was recycled by re-allocation between the two accesses.
Figure~\ref{fig:recycle} shows a common code pattern and interleaving which can expose such behavior.
If the {\tt malloc} on line 4 returns the same address passed to {\tt free} on line 2, then lines 1 and 7 will be flagged as a data race.
We call this a {\em malloc-recycle data race}.
To the human eye, this is obviously a false positive: reordering lines 4-7 before lines 1-2 will change {\tt malloc}'s return value, causing {\tt x} and {\tt y} to no longer collide.
Here, Thread 2's logic usually corresponds to an initialization pattern \cite{eraser}, but for generality we have added a {\tt publish} action on line 6.

@@

Fortunately, when data-race detection is combined with DPOR and Iterative Deepening, pruning all malloc-recycle candidates is sound, even considering adversarial programs such as Figure~\ref{fig:recycle-bug}.
This makes it unnecessary to verify such data races by actively adding more preemptions,
achieving a potentially combinatorial reduction in how many state spaces we generate.
%Intuitively, we need not worry about cases such as Figure~\ref{fig:recycle-bug} because,
%should they be true races,
%DPOR will reorder threads sufficiently for the malloc-recycle pattern to disappear.
We provide a proof sketch below.
